---
title: "ds202_lab3"
author: "Rithwik Gokhale"
date: "3/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
install.packages('readxl')
data <- readxl::read_excel('/Users/rithw/Desktop/School work/DS 202/ds202_lab3/GSS.xls')
data
```

Question 3
```{r}
happy <- filter(data, `General happiness` == "Pretty happy" | `General happiness` == "Very Happy")
ggplot(happy, aes(x = `Marital status`)) + geom_histogram(stat = "count")
data %>% 
  select(`General happiness`, `Marital status`) %>%
    group_by(`Marital status`) %>%
      count(`General happiness`)
```



We can see from the above graohs that people who are married seem to be happier than people who are widowed or have not married or chosen to remain single. The second happiest group are the not married and third is the divorced group. The least happy from the graphs are the people who have separated.

Question 4

```{r}
ggplot(data, aes(x = `Marital status`)) + geom_bar() + facet_wrap(~`Political party affiliation`) + coord_flip()
df <- data %>%
  select(`General happiness`, `Political party affiliation`, `Marital status`) %>%
    group_by(`Political party affiliation`, `Marital status`) %>%
      count(`General happiness`)
df
```



The above information shows that political affiliation does not have an effect on the relationship found in question 3. This is because the distribution of the happy couples is pretty much proportional across the different political parties. Again, regardless of the political parties, the married people are the happiest and the separated are the least happy. 

#Question 5 

How is marital status related to the education of a respondent? What are possible reasons accounting for the patterns you see?


```{r}
ggplot(data, aes(x=`Marital status`))+ 
  geom_bar() +
  facet_wrap(~ `Highest year of school completed`)+
  coord_flip()
```

It is quite interesting that there exists a pattern for the years of education through the groups by marital status. As we can see, group of married tends have more frequencies than the never married group. So, the education frequencies goes by group married > never married > divorced > widowed. There are not much of data of groups of separated, so it is hard to tell the relationship between the level of education and it. 
Looking at the facet wrap, we can conclude that the more educated the person is, the better his/her marriage life is. 