---
title: "ds202_lab3"
author: "Rithwik Gokhale and Tyren Teav"
date: "3/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1


```{r}
library(dplyr)
library(ggplot2)

install.packages('readxl', repos = "http://cran.us.r-project.org")
data <- readxl::read_excel('/Users/rithw/Desktop/School work/DS 202/ds202_lab3/GSS.xls')
data

data <- data %>% rename('year' = 'Gss year for this respondent', 'id' = `Respondent id number`, 'labor_status' = `Labor force status`, 'work_hours' = `Number of hours usually work a week`, marital_status = `Marital status`, 'years_of_school' = `Highest year of school completed`, 'political_party' = `Political party affiliation`, 'happiness' = `General happiness`, 'ballot' = `Ballot used for interview`)

unique(data$years_of_school)
years <- as.numeric(data$years_of_school)
data$years = years

data %>%
  group_by(marital_status) %>%
    summarize(
      year_avg = mean(years, na.rm = TRUE)) %>%
    arrange(year_avg)
   
```

The people who were never married had the highest average of years of school completed. There is no other real pattern, those who were widowed had the lowest average next to those who were separated. 

# Comment from rgokhale - The graphs which are used here can easily show the comparision between the level of education and the level of happiness. 

Problem 2

```{r}

widow <- data.frame(data[data$marital_status %in% c('Widowed'), ])
widow1 <- na.omit(widow)
wavg <- mean(widow1$years)
widow2 <- widow1[!widow1$years < wavg, ]

sep <- data.frame(data[data$marital_status %in% c('Separated'), ])
sep1 <- na.omit(sep)
savg <- mean(sep1$years)
sep2 <- sep1[!sep1$years < savg, ]

mar <- data.frame(data[data$marital_status %in% c('Married'), ])
mar1 <- na.omit(mar)
mavg <- mean(mar1$years)
mar2 <- mar1[!mar1$years < mavg, ]

div <- data.frame(data[data$marital_status %in% c('Divorced'), ])
div1 <- na.omit(div)
davg <- mean(div1$years)
div2 <- div1[!div1$years < davg, ]

nmar <- data.frame(data[data$marital_status %in% c('Never married'), ])
nmar1 <- na.omit(nmar)
nmavg <- mean(nmar1$years)
nmar2 <- nmar1[!nmar1$years < nmavg, ]

na <- data.frame(data[data$marital_status %in% c('No answer'), ])
na1 <- na.omit(na)
naavg <- mean(na1$years)
na2 <- na1[!na1$years < naavg, ]

educ <- rbind(widow2, sep2, mar2, div2, nmar2, na2)
educ

```

# comment from rgokhale - this code is very easy to understand and the information being displayed from the code can be used to draw straightforward conclusions

Question 3
```{r}
happy <- filter(data, `happiness` == "Pretty happy" | `happiness` == "Very Happy")
ggplot(happy, aes(x = `marital_status`)) + geom_histogram(stat = "count")
data %>% 
  select(`happiness`, `marital_status`) %>%
    group_by(`marital_status`) %>%
      count(`happiness`)
```



We can see from the above graphs that people who are married seem to be happier than people who are widowed or have not married or chosen to remain single. The second happiest group are the not married and third is the divorced group. The least happy from the graphs are the people who have separated.

Question 4

```{r}
ggplot(data, aes(x = `marital_status`)) + geom_bar() + facet_wrap(~`political_party`) + coord_flip()
df <- data %>%
  select(`happiness`, `political_party`, `marital_status`) %>%
    group_by(`political_party`, `marital_status`) %>%
      count(`happiness`)
df
```



The above information shows that political affiliation does not have an effect on the relationship found in question 3. This is because the distribution of the happy couples is pretty much proportional across the different political parties. Again, regardless of the political parties, the married people are the happiest and the separated are the least happy. 

#Question 5 

How is marital status related to the education of a respondent? What are possible reasons accounting for the patterns you see?


```{r}
ggplot(data, aes(x=`marital_status`))+ 
  geom_bar() +
  facet_wrap(~ `years_of_school`)+
  coord_flip()
```

 
From the above analysis, we can that there is a pattern between the years of education and the marital status information. So the education frequency goes by married, then never married, then divorced and then widowed. There is not sufficient information between the education level and the people who are separated. Basically, we can conclude that the more aducated a person is, the more happier their marriage life is. 

Problem 6

We will be looking at the relationship between labor status and happiness. I gave a numeric value to the happiness levels, 1 to 3, from not too happy to very happy. Then I will be taking the average to see the happiness levels of different labor statuses.The happiest were keeping the house and the unhappiest were those who were laid off or unemployed.

```{r}
library(dplyr)
library(ggplot2)
unique(data$happiness)

happynum <- as.numeric(data$happiness)

happynum = ifelse(data$happiness %in% c('Very happy'), 3, happynum)
happynum = ifelse(data$happiness %in% c('Pretty happy'), 2, happynum)
happynum = ifelse(data$happiness %in% c('Not too happy'), 1, happynum)
data$happynum = happynum


data %>% 
  group_by(labor_status) %>%
  summarize(
    avg_happy = mean(happynum, na.rm = TRUE)) %>%
      arrange(avg_happy)

happy1 <- data.frame(data[data$happynum %in% c(1), ])
happy2 <- data.frame(data[data$happynum %in% c(2), ])
happy3 <- data.frame(data[data$happynum %in% c(3), ])

happys <- rbind(data.frame(fill = 'Not too happy', hp = happy1),
                 data.frame(fill = 'Pretty happy', hp = happy2),
                 data.frame(fill = 'Very happy', hp = happy3))

ggplot(happys, aes(x=hp.labor_status, fill=fill)) + geom_bar(position = 'dodge') + coord_flip()

```

Problem 6 pt.2

We will look at years of school vs the year the survey was taken, see if the time had any affect on how much school was completed. The average of the education increased overall over the years. There are little dips, but an obvious increasing trend. It seems that the max amount did not change much, but the minimum amount increased, thus increasing the average.

# comment from rgokhale - the summary provided above connects with the data displayed from the graphs. Tyren has used easy to understand code

```{r}

b <- data %>%
  group_by(year) %>%
  summarize(
  avg = mean(years, na.rm = TRUE)) %>%
  arrange(avg)

avg = b$avg

ggplot(data, aes(x=year, y=years)) + geom_point(position = 'jitter')

ggplot(b, aes(x=year, y=avg)) + geom_point(position = 'jitter')



```

